---
title: "Project"
author: "Cheryl Machingura"
date: "2024-08-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation and Cleansing Steps

### 1. How to Import and Clean My Data

#### Importing Data
I will use R's `readr` library to import the dataset. The dataset includes information from social media usage and mental health surveys for teenagers. Here's how I will import the data:

```{r}
# Install and load necessary libraries
required_packages <- c("readr", "dplyr", "tidyr", "lubridate")
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load libraries
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)

# Importing the datasets
social_media_data <- read_csv('social_media_usage.csv')
mental_health_data <- read_csv('teen_mental_health_surveys.csv')

# Display the structure of the data
str(social_media_data)
str(mental_health_data)
```

#### Cleaning Data

Cleaning the data will involve the following steps:
1. Handling Missing Values: Identify and handle missing values by either filling them with appropriate values or dropping the rows/columns.
2. Removing Duplicates: Ensure that there are no duplicate rows.
3. Data Type Conversion: Convert columns to appropriate data types (e.g., dates, categorical variables).
4. Standardizing Column Names: Ensure column names are consistent and readable.

Cleaning Process:

```{r}
# Handling missing values
social_media_data <- social_media_data %>% drop_na()
mental_health_data <- mental_health_data %>% drop_na()

# Removing duplicates
social_media_data <- social_media_data %>% distinct()
mental_health_data <- mental_health_data %>% distinct()

# Data type conversion
social_media_data <- social_media_data %>%
  mutate(usage_date = as.Date(usage_date, format = "%Y-%m-%d"))
  
mental_health_data <- mental_health_data %>%
  mutate(survey_date = as.Date(survey_date, format = "%Y-%m-%d"))

# Standardizing column names
social_media_data <- social_media_data %>% rename_with(~ tolower(gsub(" ", "_", .)))
mental_health_data <- mental_health_data %>% rename_with(~ tolower(gsub(" ", "_", .)))
```

### 2. Final Data Set

Condensed view of the final dataset after cleaning:

#### Social Media Data (First 5 rows)

```{r}
head(social_media_data, 5)
```

#### Mental Health Data (First 5 rows)

```{r}
head(mental_health_data, 5)
```

### 3. Questions for Future Steps

1. How to effectively merge these datasets based on user_id and date?
2. What additional cleaning might be necessary after merging?
3. What specific statistical methods or machine learning techniques should be used for deeper analysis?

### 4. What Information Is Not Self-Evident?

The correlation between social media usage and mental health indicators is not immediately evident. The data needs to be analyzed to uncover patterns, trends, and potential causations.

### 5. Different Ways to Look at This Data

- Time Series Analysis: Analyzing changes over time.
- Correlation Analysis: Checking for correlations between social media usage and mental health scores.
- Categorical Analysis: Comparing mental health scores across different social media platforms.

### 6. How Data will be Sliced and Diced

- By Platform: Comparing mental health scores among users of different social media platforms.
- By Time: Observing changes in mental health scores over different time periods.
- By Usage: Grouping data based on hours spent on social media to see its effect on mental health.

### 7. Summarizing Data to Answer Key Questions

- Descriptive Statistics: Mean, median, standard deviation of usage hours and mental health scores.
- Correlation Coefficients: To determine the strength and direction of relationships between variables.

### 8. Plots and Tables to Illustrate the Findings

- Line Plots: For time series analysis.
- Bar Charts: For comparing different platforms.
- Scatter Plots: For showing correlations between usage hours and mental health scores.
- Box Plots: For visualizing distributions and detecting outliers.

### 9. Incorporating Machine Learning Techniques

I plan to use regression analysis to predict mental health scores based on social media usage patterns. Clustering techniques may also be used to identify distinct groups within the data.

### 10. Questions for Future Steps

1. How to interpret and validate the results from machine learning models?
2. What additional features or data sources could improve the analysis?
3. How to effectively communicate findings to a non-technical audience?
```

